The Neuro Bureau ADHD-200 Preprocessed Repository

### 2015-12-21 Bellec_ADHD200_2016_backup.tex
Re-organized derivatives and pipelines. Added substantial detail to describtions of Athena and Burner pipelines.

%%
%% Start line numbering here if you want
%%
% \linenumbers
\section{Highlights}
\begin{itemize}
\item Processed derivatives generated from structural and resting-state fMRI in 973 individuals. 
\item The sample includes 585 typically developing children, 362 children with attention deficit hyperactivity disorder, and 26 with unknown diagnostic status.
\item Functional MRI volumes were denoised and aligned using two alternative processing pipelines. 
\item Derivatives include regional time series, grey matter segmentation as well as two variants of connectivity maps and a measure of intrinsic brain activity. 
\item These resources are freely available, with minimal access requirements. 
\end{itemize}
\section{Introduction}

In 2011, the ``ADHD-200 Global Competition'' was held with the aim of engaging researchers from a variety of analytical backgrounds to identify biomarkers of attention deficit hyperactivity disorder (ADHD) from resting-state functional magnetic resonance imaging (rs-fMRI) and structural MRI (s-MRI) data \cite{Milham2012}. The competition made use of the ``ADHD-200 Sample'' data collection that was aggregated from eight independent sites and shared through the Intenational Neuroimaging Datasharing Initiative (INDI) \cite{Mennes2013}. The data includes rs-fMRI, structural MRI (sMRI), and basic phenotypic information for 973 individuals (585 TDC, 362 ADHD, 26 unknown) \cite{Milham2012}. Competitors were given five and a half months to optimize a classification algorithm on training data (776 individuals) and submit their predicted clinical labels for test data for which diagnostic information was withheld. The competition data was distributed in a raw form and before any analysis could begin, had to be preprocessed to make it comparable across individuals and to remove nuisance variation that would obscure the underlying biological signal. These preprocessing steps present a significant hurdle for would-be competitors who do not have the specialist knowledge of neuroimaging methods or access to the high performance computing resources required to process such a large data collection in a short amount of time. This category includes statisticians and computer scientists, who were potentially the most qualified for the machine learning aspect of the competition. Realizing this barrier to entry, the Neuro Bureau prospectively collaborated with all competitors by preprocessing the data and sharing these results.

The ``ADHD-200 Preprocessed'' is a repository of preprocessed rs-fMRI and s-MRI data along with statistical derivatives from the ADHD-200 Sample. Rather than favoring a specific processing strategy, we followed a pluralistic approach by preprocessing the data using multiple pipelines (called ``Athena'', ``Burner'' and ``NIAK'') that differed in toolsets used, the philosophy motivating choices of algorithms and parameters, and the statistical derivatives calculated. The Athena pipeline processed rs-fMRI and s-MRI images using a custom Bash script that combined AFNI \cite{cox1996afni} and FSL \cite{smith2004advances} neuroimaging toolkits. The Burner pipeline used SPM8 \cite{ashburner2012spm8} to process s-MRI data for voxel-based morphometry style tools. The NIAK pipeline processed rs-fMRI and s-MRI using tools from the neuroimaging analysis kit (NIAK) \cite{Bellec2011} assembled together into a workflow using the Pipeline System for Octave and Matlab (PSOM) \cite{Bellec2012}. The results from the pipelines are shared on the Neuro Bureau project page at the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC)\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. The data was used by several teams who competed in the ADHD-200 Global Competition, including the winning entry by a group of biostaticians with limited prior neuroimaging experience \cite{Eloyan2012}. The initiative has had an significant impact beyond the competition, having been used in 49 publications \cite{Rangarajan2014, Liang2012, Tabas2014, Rangarajan2015, Mahanand2013, Lifshitz2012, Fujita2013, Ji2011, Li2015, Li2013, Liu2012, DosSantosSiqueira2014, Olivetti2014, Han2015, Wang2013a, Subramanian2013, Dey2014, Bellec2012, Bohland2012, Chang2012, Cheng2012, Colby2012, Dai2012, Dey2012, Eloyan2012, Olivetti2012, Sato2012a, Carmona2015, Carmona2015a, Hou2015, Deshpande2015, She2014, Lavoie-Courchesne2012b, Chen2015, Nunez-Garcia2015, Solmaz2012, Anderson2014, KadkhodaeianBakhtiari2012, Sato2013, Kyeong2015, Sato2012, Takahashi2012, He2013, Kong2013, Yao2013, Yang2015, Ahn2015, Fujita2014, Reiss2014}, three PhD theses \cite{Colby2012a, Dey2013, Zhang2012}, three master's dissertations \cite{VanGalenLast2011, Vidal2014, Wang2013} and one patent \cite{Dey2013} in just over 3 years, with more in press or under review.

@@

The ADHD-200 preprocessed data was released in the summer of 2012 and can be downloaded from NITRC\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. No data usage agreement is required to access or download the data, the only requirement is registering for a free NITRC account. This registration enables downloads to be tracked for usage statistics and so that users can be contacted in the event that errors are found. The ADHD-200 Sample allows unrestricted data usage for non-commercial research purposes provided that the specific datasets included in an analysis be cited appropriately and that their funding sources be acknowledged\footnote{\url{http://fcon_1000.projects.nitrc.org/indi/adhd200/}}. There are no more restrictions placed on the preprocessed data or derivatives other than the request that the ADHD-200 Preprocessed Initiative is cited appropriately and that the specific pipeline is acknowledged in publications using the data. A forum is available on the Neuro Bureau's NITRC project page for users to ask questions or report
problems\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=2046}}. Questions regarding data acquisition or phenotypic variables should be directed to INDI's support forum\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=1735}}.

@@

ADHD-200 Preprocessed data is available as a collection of compressed tar files that are organized by pipeline, sites of data collection, training and test samples, as well as by derivative. The following preprocessed derivatives are available.

\subsection{Preprocessed s-MRI and rs-fMRI} Both the Athena and NIAK pipelines released denoised rs-fMRI volumes resampled into a reference, stereotaxic space (variants of the MNI template) as compressed 4D nifti files (.nii.gz). Transformations from native to stereotaxic space for each dataset were released in separate files, in FSL .xfm transform files for Athena and in .xfm MINC files for NIAK. In addition, the estimated motion parameters (bundled with additional intermediate outputs) were released in separate archives, inside text files (AFNI .1D tab-separated-value files) for Athena, and inside Matlab HDF5 files for NIAK.

\subsection{Regional time series} Regional time series were extracted based on a series of different brain parcellations. The Athena ppipeline used functional parcels generated by a spectral clustering method (Craddock et al. 2012) as well as a number of parcellations established from anatomical landmarks, with a resolution ranging from 93 to 351 parcels. The NIAK pipeline only used functional parcels generated by a region growing method (Bellec et al. 2006) with two resolutions: 954 and 2843 parcels. The Athena pipeline distributed regional time series as text files (.1D AFNI) and the NIAK distributed them as HDF5 files (.mat Matlab). All brain parcellations were also released as compressed nifti files (.nii.gz). Note that the phenotypic information of the ADHD200 sample was also released as comma-separated-values files (.csv).

\subsection{Connectivity maps}
The Athena pipeline derived a number of brain maps capturing some aspects of RfMRI connectivity, derived from the preprocessed fMRI datasets: fractional amplitude of low frequency fluctuations (fALFF) maps, regional homogeneity (ReHO) maps and dual-regression based on the (Smith et al. 2009) group independent component analysis (ICA) with 10 resting-state networks. One compressed nifti file (.nii.gz) was released per fMRI dataset and type of derivative, with in addition the time courses of the dual regression available in text files (.1D AFNI).

@@

\begin{sidewaystable}[!ht]
\caption{{\bf Resting state fMRI acquisition parameters by site.} Seq: imaging sequence, FA: flip angle, TE: echo time, TR: repetition time, PA: parallel acquisition, N$_{slc}$: number of slices, Th.: slice thickness, Slc. Acq.: slice acquisition order, N$_{TR}$: number of measurements (TRs),  BHBU: Bradley Hospital/ Brown University, KKI: Kennedy Krieger Institute, NI: NeuroIMAGE sample, NYU: New York University Child Study Center, OHSU: Oregon Health Sciences University, PKU: Peking University, Pitt: University of Pittsburgh, Pitt 2: U. Pitt. parameters used for acquiring the testing data, WUSTL: Washington University at Saint Louis, EPI: echo planar imaging, PACE: Prospective Acquisition CorrEction (EPI with prospective motion correction), S: sensitivity encoding (SENSE), G: generalized autocalibrating partially parallel acquisition (GRAPPA), int+: slices were acquired interleaved ascending, seq+: slices were acquired sequentially ascending, var.: the number of measurements varies across datasets, fixate: participants were asked to keep their eyes open and fixate on an image, closed: participants were asked to keep their eyes closed, open: participants were asked to keep their eyes open.}\label{smri_params_table}
  \begin{tabular}{llllllllllll}
    \hline
    Site & Seq & FA & TE & TR & PA & N$_{slc}$ & Th. & Slc. Acq. & Resolution & N$_{TR}$ & Instructions  \\ 
        \hline
    \noalign{\vskip 1ex} 
    BHBU & PACE & 90$^{\circ}$ & 25 ms & 2000 ms  & None & 35 & 3 mm & int+ & $3.0 \times 3.0$ mm$^{2}$ & 256 & fixate \\
    KKI & EPI & 75$^{\circ}$ & 30 ms & 2500 ms  & S $\times3$ & 47 & 3 mm & seq+ & $3.0 \times 3.0$ mm$^{2}$ & 128 & fixate \\
    NI & EPI & 80$^{\circ}$ & 40 ms & 1960 ms  & G $\times2$ & 37 & 3 mm & int+ & $3.5 \times 3.5$ mm$^{2}$ & 266 & eyes closed \\
    NYU & EPI & 90$^{\circ}$ & 15 ms & 2000 ms  & None & 33 & 4 mm & int+ & $3.0 \times 3.0$ mm$^{2}$ & 180 & eyes closed \\
    OHSU & EPI & 90$^{\circ}$ & 30 ms & 2500 ms  & None & 36 & 3.8 mm & int+ & $3.8 \times 3.8$ mm$^{2}$ & 82 & fixate \\
    PKU 1 & EPI & 90$^{\circ}$ & 30 ms & 2000 ms  & None & 33 & 4.2 mm & int+ & $3.1 \times 3.1$ mm$^{2}$ & 240 & closed or fixate \\
    PKU 2 & EPI & 90$^{\circ}$ & 30 ms & 2000 ms  & None & 33 & 3.6 mm & int+ & $3.1 \times 3.1$ mm$^{2}$ & 240 & closed or fixate \\
    PKU 3 & EPI & 90$^{\circ}$ & 30 ms & 2000 ms  & None & 30 & 4.5 mm & int+ & $3.44 \times 3.44$ mm$^{2}$ & 240 & closed or fixate \\
    Pitt & EPI & 70$^{\circ}$ & 29 ms & 1500 ms  & G $\times2$ & 29 & 4.0 mm & int+ & $3.1 \times 3.1$ mm$^{2}$ & 200 & open or closed \\
    Pitt 2 & EPI & 90$^{\circ}$ & 30 ms & 3000 ms  & None & 46 & 3.5 mm & int+ & $3.8 \times 3.8$ mm$^{2}$ & 128 & open or closed \\
    WUSTL & EPI & 90$^{\circ}$ & 27 ms & 2500 ms  & None & 32 & 4.0 mm & int+ & $4.0 \times 4.0$ mm$^{2}$ & var. & fixate \\
    \noalign{\vskip 1ex} 
    \hline

@@

\subsection{Preprocessed sMRI and rs-fMRI}
% http://onlinelibrary.wiley.com/doi/10.1002/hbm.22790/epdf
\paragraph{The Athena pipeline} The preprocessing steps of functional MRI datasets in the Athena pipeline included the following steps. (1) Removal of the first four volumes to allow for magnetization to reach equilibrium. (2) Slice timing correction. (3) Motion correction by means of estimation of a rigid body transform for each time frame. (4) Normalization to MNI152 stereotaxic space. (5) Resampling at 4 mm isotropic voxel resolution. (6) Regression of nuisance covariates including head motion parameters, global mean, white matter, and cerebrospinal ﬂuid signals. (7) Band-pass filtering of fMRI ($0.009 Hz < f < 0.08 Hz$) to reduce low-frequency drift and high-frequency physiological noise. (8) Spatial smoothing with a 6 mm full-width-at-half-maximum isotropic Gaussian kernel.

\paragraph{The NIAK pipeline} The preprocessing steps in NIAK followed those of the Athena pipeline with differences at the following steps. (1) Only the first three volumes were suppressed, instead of four. (4) Although both Athena and NIAK used the MNI non-linear stereotaxic template space described in \citep{fonov2011unbiased}, the NIAK used the young adult (ICBM152) symmetric variant, while Athena used the NIHPD asymmetric pediatric template (4.5 y.o.-18.5 y.o.). (5) The fMRI volumes were resampled in stereotaxic space at 3 mm isotropic (instead of 4 mm), and this was done only right before spatial smoothing (step 5 was moved before step 8). (6) No regression of confounds was performed. Instead, a physiological noise correction based on automatic labeling of ICA components was implemented \citep{perlbarg2007corsica} after temporal filtering (steps 6 and 7 were inverted). (7) Only a high-pass filter was implemented ($0.009 Hz < f$), instead of a band pass.

@@

\paragraph{Anatomical parcellations} For both Athena and NIAK, regional time series were extracted by averaging preprocessed resting-state volumes within different brain parcels. The Athena pipeline first used a series of parcellations delineated on the basis of anatomical landmarks:
\begin{itemize}
\item The automated anatomical labeling (AAL) template \citep{tzourio2002automated}, as distributed with the AAL Toolbox. 
\item The Talairach and Tournoux (TT) parcellation \citep{Lancaster2000automated}. The TT atlas distributed with AFNI was coregistered and warped into template space.
\item The Eickhoff-Zilles (EZ) parcellation \citep{eickhoff2005new}. The EZ atlas was derived from the max-propagation atlas distributed with the SPM Anatomy Toolbox. The atlas was transformed into template space using the Colin 27 template (also distributed with the toolbox) as an intermediary. 
\item The Harvard-Oxford (HO) parcellation, distributed with FSL, was split into cortical and subcortical probabilistic atlases. A 25\% threshold was applied to each of these atlases and they were subsequently bisected into left and right hemispheres at the midline ($x=0$). ROIs representing left/right WM, left/right GM, left/right CSF and brainstem were removed from the subcortical atlas. The subcortical and cortical ROIs were combined.
\end{itemize}
All the anatomical templates were resampled into the functional space using nearest-neighbor interpolation.

@@

\subsection{Connectivity maps}
% We are missing the ALFF
Regional homogeneity (ReHo) was calculated for the preprocessed and filtered whole brain rs-fMRI data using the procedure described in \citep{zang2004regional}. The functional connectivity maps for 10 resting-state networks reported by \cite{smith2009correspondence} were constructed using a modified dual regression approach. A spatial multiple regression was first used to extract time courses corresponding to each network. In a second step, each time course was independently correlated with whole brain time series to generate subject-specific functional connectivity maps for each network. Alternatively, all time courses were entered simultaneously into a multiple (temporal) regression, and the regression cofficients associated with each time course constituted the functional connectivity maps. Both approaches have been released.

% The scripts used to calculate ReHo were generously provided by Dr. Xinian Zuo.
\subsection{Voxel-based morphometry}
%Details on the Burner pipeline can be found on the NITRC website\footnote{\url{http://www.nitrc.org/plugins/mwiki/index.php/neurobureau:NiakPipeline}}.
The T1 images were segmented into grey matter and white matter probability maps using the "New Segmentation" method implemented in SPM8. The tissue maps were rigidly aligned (translation+rotation). Inter subject normalization was performed using DARTEL toolbox in SPM8. Images were iteratively registered to the group average (population template), and the template was updated iteratively. This iterative process resulted into finer and finer registrations. The registration parameters were applied to each image to resample them into the space of the population average. A modulation was applied to preserve the global tissue volumes after normalization.

>>>>

%%
%% Start line numbering here if you want
%%
% \linenumbers
\section*{Highlights}
\begin{itemize}
\item Processed derivatives generated from structural and resting-state fMRI in 973 individuals. 
\item The sample includes 585 typically developing children, 362 children with attention deficit hyperactivity disorder, and 26 with unknown diagnostic status.
\item Functional MRI volumes were denoised and aligned using two alternative processing pipelines. 
\item Derivatives include regional time series, grey matter segmentation as well as two variants of connectivity maps and a measure of intrinsic brain activity. 
\item These resources are freely available, with minimal access requirements. 
\end{itemize}
\section{Introduction}

In 2011, the ``ADHD-200 Global Competition'' was held with the aim of engaging researchers from a variety of analytical backgrounds to identify biomarkers of attention deficit hyperactivity disorder (ADHD) from resting-state functional magnetic resonance imaging (rs-fMRI) and structural MRI (s-MRI) data \cite{Milham2012}. The competition made use of the ``ADHD-200 Sample'' data collection that was aggregated from eight independent sites and shared through the Intenational Neuroimaging Datasharing Initiative (INDI) \cite{Mennes2013}. The data includes rs-fMRI, structural MRI (sMRI), and basic phenotypic information for 973 individuals (585 TDC, 362 ADHD, 26 unknown) \cite{Milham2012}. Competitors were given five and a half months to optimize a classification algorithm on training data (776 individuals) and submit their predicted clinical labels for test data for which diagnostic information was withheld. The competition data was distributed in a raw form and, before any analysis could begin, the images had to be preprocessed to make them comparable across individuals and reduce noise. These preprocessing steps present a significant hurdle for would-be competitors who do not have the specialist knowledge of neuroimaging methods or access to high performance computing resources. Realizing this barrier to entry, the Neuro Bureau prospectively collaborated with all competitors by preprocessing the data and sharing these results.

The ``ADHD-200 Preprocessed'' is a repository of preprocessed rs-fMRI and s-MRI data along with statistical derivatives from the ADHD-200 Sample. Rather than favoring a specific processing strategy, we followed a pluralistic approach by preprocessing the data using multiple pipelines (called ``Athena'', ``Burner'' and ``NIAK'') that differed in toolsets used, the philosophy motivating choices of algorithms and parameters, and the statistical derivatives calculated. The Athena pipeline processed rs-fMRI and s-MRI images using a custom Bash script that combined AFNI \cite{cox1996afni} and FSL \cite{smith2004advances} neuroimaging toolkits. The Burner pipeline used SPM8 \cite{ashburner2012spm8} to process s-MRI data for voxel-based morphometry style tools. The NIAK pipeline processed rs-fMRI and s-MRI using a workflow composed of tools from the neuroimaging analysis kit (NIAK) \cite{Bellec2011}, which were assembled with the Pipeline System for Octave and Matlab (PSOM) \cite{Bellec2012}.

% PB -> CC: the following information is redundant with the rest of the paper. It's also not introductory material. As the paper is short, I would tend to avoid repetition.

%The results from the pipelines are shared on the Neuro Bureau project page at the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC)\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. The data was used by several teams who competed in the ADHD-200 Global Competition, including the winning entry by a group of biostaticians with limited prior neuroimaging experience \cite{Eloyan2012}. The initiative has had an significant impact beyond the competition, having been used in 49 publications \cite{Rangarajan2014, Liang2012, Tabas2014, Rangarajan2015, Mahanand2013, Lifshitz2012, Fujita2013, Ji2011, Li2015, Li2013, Liu2012, DosSantosSiqueira2014, Olivetti2014, Han2015, Wang2013a, Subramanian2013, Dey2014, Bellec2012, Bohland2012, Chang2012, Cheng2012, Colby2012, Dai2012, Dey2012, Eloyan2012, Olivetti2012, Sato2012a, Carmona2015, Carmona2015a, Hou2015, Deshpande2015, She2014, Lavoie-Courchesne2012b, Chen2015, Nunez-Garcia2015, Solmaz2012, Anderson2014, KadkhodaeianBakhtiari2012, Sato2013, Kyeong2015, Sato2012, Takahashi2012, He2013, Kong2013, Yao2013, Yang2015, Ahn2015, Fujita2014, Reiss2014}, three PhD theses \cite{Colby2012a, Dey2013, Zhang2012}, three master's dissertations \cite{VanGalenLast2011, Vidal2014, Wang2013} and one patent \cite{Dey2013} in just over 3 years, with more in press or under review.

@@

The ADHD-200 preprocessed data was released in 2011 and can be downloaded from NITRC\footnote{\url{http://www.nitrc.org/frs/?group_id=383}}. No data usage agreement is required to access or download the data, the only requirement is registering for a free NITRC account. This registration enables downloads to be tracked for usage statistics. Registered users can also be contacted in the event that errors would be found. The ADHD-200 Sample allows unrestricted data usage for non-commercial research purposes provided that the specific datasets included in an analysis be cited appropriately and that their funding sources be acknowledged\footnote{\url{http://fcon_1000.projects.nitrc.org/indi/adhd200/}}. There are no more restrictions placed on the preprocessed data or derivatives other than the request that the ADHD-200 Preprocessed Initiative is cited appropriately and that the specific pipeline is acknowledged in publications using the data. A forum is available on the Neuro Bureau's NITRC project page for users to ask questions or report
problems\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=2046}}. Questions regarding data acquisition or phenotypic variables should be directed to INDI's support forum\footnote{\url{http://www.nitrc.org/forum/forum.php?forum_id=1735}}.

@@

The ADHD-200 Preprocessed repository contains preprocessed outputs and derivatives for data from the ADHD-200 Sample, which includes 973 individuals (352 F) between the ages of 7 and 27 aggregated from 17 different studies conducted across 8 different sites (for a breakdown of age and sex by diagnosis see table \ref{part_table}). For each individual, phenotypic data includes sex, age, handedness, ADHD diagnosis (585 TDC, 362 ADHD, 26 unknown), ADHD subtype (ADHD-combined, ADHD-inattentive, ADHD-hyperactive/impulsive), one of three different measures of ADHD severity, one of five measures of intelligence, co-morbid diagnoses, and whether or not they have used medication to treat their symptoms \cite{Milham2012}. Imaging data for each individual includes one or more T1-weighted high-resolution s-MRI scan (s-MRI) and one or more rs-fMRI scan. Most of the data was collected from a single imaging session, although a second session is available for 15 individuals from the WUSTL site. There is a substantial amount of variation in how the data was collected across sites including the type of MRI system and scanning parameters used, the length of the rs-fMRI scans, and the instructions given to the participants prior to the scan (see tables \ref{smri_params_table} and \ref{rsfmri_params_table}).

Nearly all of the imaging data from the ADHD-200 Sample was included in the preprocessing effort, some individuals were excluded for poor quality or missing data. The results of the preprocessing are made available as a collection of compressed tar files that are organized by pipeline, sites of data collection, training and test samples, as well as by derivative. A group level file containing this phenotypic data is available in comma-separated-values format (.csv).

% PB->CC I did the breakdown by derivatives, not by pipeline. Otherwise there is too much repetition between NIAK and Athena. This is true for content description but even more for the methods section, where it's nice for example to have the different approaches for generating parcellations under one section, rather than disjointed into two "Athena" and "NIAK" sections. I've also left the details of the methods (e.g. resolution for resampling) to the methods section.

% I understand, but I am not sure that the repetition is bad. From my understanding, since this is essentially a data paper, we need to be very explicit about what data is available. the way it was before missed alot of useful details, i haven't removed your original content yet, just experimenting with an alternative approach. I would actually prefer it if we could put it into a table, but since what is available varies so much by pipeline, it would be a very lopsided table. Also, since most of the derivatives are pipeline specific, you are essentially breaking them down by derivatives. I think it would be better if the processing and 'contents' sections were merged.

\subsection{Athena Pipeline} Data from the Athena pipeline includes processed structural data and processed functional data along with extracted features and statistical derivatives. For the structural data the repository contains the skull-stripped whole brain images, along with spatially smoothed and unsmoothed grey matter density maps, written into template space as compressed 3D nifti files (nii.gz). The transformation from native space to the template is provided as a FSL fNIRT warp file (nii.gz). For the functional data, the repository contains denoised rs-fMRI volumes, processed with and without temporal bandpass filtering, written into template space at 4 $\times$ 4 $\times$ 4 mm$^3$ resolution and spatially smoothed with a 6-mm FWHM Gaussian filter (compressed 4D nifti, nii.gz). Ancillary files including the mean rs-fMRI image and masks that differentiate brain data from background in template space (nii.gz), and six parameter head motion traces (tab separated value, AFNI .1D files). Regional time series extracted using four different anatomical atlases\cite{tzourio2002automated, lancaster2000automated, eickhoff2005new, HO_atlas_1, HO_atlas_2, HO_atlas_3, HO_atlas_4}, two atlases derived by clustering the functional data \cite{craddock2012whole}, and 10 highly reproducible intrinsic connectivity networks (ICNs) \cite{smith2009correspondence}  are available in tab separated value files (AFNI .1D). Derivatives include fractional amplitude of low frequency fluctuations (fALFF) \cite{zuo2008falff},  spatial maps for the 10 ICNs calculated from the non-bandpass-filtered data, and regional homogeneity (ReHo) \cite{zang2004regional} maps calculated from the bandpass-filtered data. The derivatives are available in template space at 4 $\times$ 4 $\times$ 4 mm$^3$ resolution and smoothed with a 6-mm FWHM Gaussian filter as compressed 3D (fALFF, ReHo) or 4D nifti (ICNs) files (.nii.gz).

\subsection{Preprocessed s-MRI and rs-fMRI} Both the Athena and NIAK pipelines released denoised rs-fMRI volumes resampled into a reference, stereotaxic space (variants of the MNI template) as compressed 4D nifti files (.nii.gz). Transformations from native to stereotaxic space for each dataset were released in separate files, in FSL wrap .nii.gz file for Athena and in .xfm/.mnc MINC wrap files for NIAK. In addition, ancillary files such as the estimated rigid-body motion parameters (AFNI .1D tab-separated-value for Athena, and HDF5 files .mat files for NIAK), mean functional image (.nii.gz files) as well as individual and group brain masks (.nii.gz files) were released in separate archives.

\subsection{Regional time series} Regional time series were extracted based on a series of different brain parcellations. The Athena ppipeline used functional parcels generated by a spectral clustering method \cite{craddock2012whole} as well as a number of parcellations established from anatomical landmarks, with a resolution ranging from 93 to 351 parcels. The NIAK pipeline only used functional parcels generated by a region growing method \cite{bellec2006identification} with two resolutions: 954 and 2843 parcels. The Athena pipeline distributed regional time series as text files (.1D AFNI) and the NIAK distributed them as HDF5 files (.mat Matlab). All brain parcellations were also released as compressed nifti files (.nii.gz). Note that the phenotypic information of the ADHD200 sample was also released as comma-separated-values files (.csv).

\subsection{Brain maps}
The Athena pipeline derived a number of brain maps derived from the preprocessed fMRI datasets: fractional amplitude of low frequency fluctuations (fALFF) maps, regional homogeneity (ReHO) maps and dual-regression based on a group independent component analysis (ICA) with 10 resting-state networks \citep{smith2009correspondence}. One compressed nifti file (.nii.gz) was released per fMRI dataset and type of brain map, with in addition the time courses of the dual regression available in text files (.1D AFNI).

@@

\begin{sidewaystable}[!ht]
\caption{{\bf Resting state fMRI acquisition parameters by site.} Seq: imaging sequence, FA: flip angle, TE: echo time, TR: repetition time, PA: parallel acquisition, N$_{slc}$: number of slices, Th.: slice thickness, Slc. Acq.: slice acquisition order, N$_{TR}$: number of measurements (TRs),  BHBU: Bradley Hospital/ Brown University, KKI: Kennedy Krieger Institute, NI: NeuroIMAGE sample, NYU: New York University Child Study Center, OHSU: Oregon Health Sciences University, PKU: Peking University, Pitt: University of Pittsburgh, Pitt 2: U. Pitt. parameters used for acquiring the testing data, WUSTL: Washington University at Saint Louis, EPI: echo planar imaging, PACE: Prospective Acquisition CorrEction (EPI with prospective motion correction), S: sensitivity encoding (SENSE), G: generalized autocalibrating partially parallel acquisition (GRAPPA), int+: slices were acquired interleaved ascending, seq+: slices were acquired sequentially ascending, var.: the number of measurements varies across datasets, fixate: participants were asked to keep their eyes open and fixate on an image, closed: participants were asked to keep their eyes closed, open: participants were asked to keep their eyes open.}\label{rsfmri_params_table}
  \begin{tabular}{llllllllllll}
    \hline
    Site & Seq & FA & TE & TR & PA & N$_{slc}$ & Th. & Slc. Acq. & Resolution & N$_{TR}$ & Instructions  \\ 
        \hline
    \noalign{\vskip 1ex} 
    BHBU & PACE & 90$^{\circ}$ & 25 ms & 2000 ms  & None & 35 & 3 mm & int+ & $3.0 \times 3.0$ mm$^{2}$ & 256 & fixate \\
    KKI & EPI & 75$^{\circ}$ & 30 ms & 2500 ms  & S $\times3$ & 47 & 3 mm & seq+ & $3.0 \times 3.0$ mm$^{2}$ & 128 & fixate \\
    NI & EPI & 80$^{\circ}$ & 40 ms & 1960 ms  & G $\times2$ & 37 & 3 mm & int+ & $3.5 \times 3.5$ mm$^{2}$ & 266 & eyes closed \\
    NYU & EPI & 90$^{\circ}$ & 15 ms & 2000 ms  & None & 33 & 4 mm & int+ & $3.0 \times 3.0$ mm$^{2}$ & 180 & eyes closed \\
    OHSU & EPI & 90$^{\circ}$ & 30 ms & 2500 ms  & None & 36 & 3.8 mm & int+ & $3.8 \times 3.8$ mm$^{2}$ & 82 & fixate \\
    PKU 1 & EPI & 90$^{\circ}$ & 30 ms & 2000 ms  & None & 33 & 4.2 mm & int+ & $3.1 \times 3.1$ mm$^{2}$ & 240 & closed or fixate \\
    PKU 2 & EPI & 90$^{\circ}$ & 30 ms & 2000 ms  & None & 33 & 3.6 mm & int+ & $3.1 \times 3.1$ mm$^{2}$ & 240 & closed or fixate \\
    PKU 3 & EPI & 90$^{\circ}$ & 30 ms & 2000 ms  & None & 30 & 4.5 mm & int+ & $3.44 \times 3.44$ mm$^{2}$ & 240 & closed or fixate \\
    Pitt & EPI & 70$^{\circ}$ & 29 ms & 1500 ms  & G $\times2$ & 29 & 4.0 mm & int+ & $3.1 \times 3.1$ mm$^{2}$ & 200 & open or closed \\
    Pitt 2 & EPI & 90$^{\circ}$ & 30 ms & 3000 ms  & None & 46 & 3.5 mm & int+ & $3.8 \times 3.8$ mm$^{2}$ & 128 & open or closed \\
    WUSTL & EPI & 90$^{\circ}$ & 27 ms & 2500 ms  & None & 32 & 4.0 mm & int+ & $4.0 \times 4.0$ mm$^{2}$ & var. & fixate \\
    \noalign{\vskip 1ex} 
    \hline

@@

\subsection{Preprocessed sMRI and rs-fMRI}
% http://onlinelibrary.wiley.com/doi/10.1002/hbm.22790/epdf
\paragraph{The Athena pipeline} The preprocessing steps of functional MRI datasets in the Athena pipeline included the following steps. (1) Removal of the first four volumes to allow for magnetization to reach equilibrium. (2) Slice timing correction. (3) Motion correction by means of estimation of a rigid body transform for each time frame. (4) Normalization to MNI152 stereotaxic space. (5) Resampling at 4 mm isotropic voxel resolution. (6) Regression of nuisance covariates including head motion parameters, global mean, white matter, and cerebrospinal ﬂuid signals. (7) Band-pass filtering of fMRI ($0.009$ Hz $< f < 0.08$ Hz) to reduce low-frequency drift and high-frequency physiological noise. (8) Spatial smoothing with a 6 mm full-width-at-half-maximum isotropic Gaussian kernel.

\paragraph{The NIAK pipeline} The preprocessing steps in NIAK followed those of the Athena pipeline with differences at the following steps. (1) Only the first three volumes were suppressed, instead of four. (4) Although both Athena and NIAK used the MNI non-linear stereotaxic template space described in \citep{fonov2011unbiased}, the NIAK used the young adult (ICBM152) symmetric variant, while Athena used the NIHPD asymmetric pediatric template (4.5 y.o.-18.5 y.o.). (5) The fMRI volumes were resampled in stereotaxic space at 3 mm isotropic (instead of 4 mm), and this was done only right before spatial smoothing (step 5 was moved before step 8). (6) No regression of confounds was performed. Instead, a physiological noise correction based on automatic labeling of ICA components was implemented \citep{perlbarg2007corsica} after temporal filtering (steps 6 and 7 were inverted). (7) Only a high-pass filter was implemented ($0.01$ Hz $< f$), instead of a band pass.

@@

\paragraph{Anatomical parcellations} For both Athena and NIAK, regional time series were extracted by averaging preprocessed resting-state volumes within different brain parcels. The Athena pipeline first used a series of parcellations delineated on the basis of anatomical landmarks:
\begin{itemize}
\item The automated anatomical labeling (AAL) template \citep{tzourio2002automated}, as distributed with the AAL Toolbox. 
\item The Eickhoff-Zilles (EZ) parcellation \citep{eickhoff2005new}. The EZ atlas was derived from the max-propagation atlas distributed with the SPM Anatomy Toolbox. The atlas was transformed into template space using the Colin 27 template (also distributed with the toolbox) as an intermediary. 
\item The Harvard-Oxford (HO) parcellation, distributed with FSL, was split into cortical and subcortical probabilistic atlases. A 25\% threshold was applied to each of these atlases and they were subsequently bisected into left and right hemispheres at the midline ($x=0$). ROIs representing left/right WM, left/right GM, left/right CSF and brainstem were removed from the subcortical atlas. The subcortical and cortical ROIs were combined.
\item The Talairach and Tournoux (TT) parcellation \citep{Lancaster2000automated}. The TT atlas distributed with AFNI was coregistered and warped into template space.
\end{itemize}
All the anatomical templates were resampled into the functional space using nearest-neighbor interpolation.

@@

\subsection{Brain maps}
% We are missing the ALFF
Regional homogeneity (ReHo) was calculated for the preprocessed and filtered whole brain rs-fMRI data using the procedure described in \citep{zang2004regional}. The functional connectivity maps for 10 resting-state networks reported by \cite{smith2009correspondence} were constructed using a modified dual regression approach. A spatial multiple regression was first used to extract time courses corresponding to each network. In a second step, each time course was independently correlated with whole brain time series to generate subject-specific functional connectivity maps for each network. Alternatively, all time courses were entered simultaneously into a multiple (temporal) regression, and the regression cofficients associated with each time course constituted the functional connectivity maps. Both approaches have been released.

% The scripts used to calculate ReHo were generously provided by Dr. Xinian Zuo.
\subsection{Voxel-based morphometry}
%Details on the Burner pipeline can be found on the NITRC website\footnote{\url{http://www.nitrc.org/plugins/mwiki/index.php/neurobureau:NiakPipeline}}.
The T1 images were segmented into grey matter and white matter probability maps using the "New Segmentation" method implemented in SPM8. The tissue maps were rigidly aligned (translation and rotation). Inter subject normalization was performed using the DARTEL toolbox in SPM8. Images were iteratively registered to the group average (population template), and the template was updated iteratively. The registration parameters were applied to each image to resample them into the space of the population average. A modulation was applied to preserve the global tissue volumes after normalization.
